---
title: "My package-building workflow"
description: |
  How I build packages 
author: Nathan L. Brouwer
date: 2022-05-25
output:
  distill::distill_article:
    self_contained: false
editor_options: 
  chunk_output_type: console
draft: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

I principally use packages to bundle up datasets and small functions for using while teaching.

The following code outlines the key steps I use to set up the framework for a package, principally using the `usethis` package.  I build a website for the package using `packagedown`.


## Packages

I use these packages for setting things up

```{r, eval = F}
library(devtools)
library(usethis)
library(here)
library(packagedown)
```

## Create repository usig GitHub

I create the git repo via the GitHub website:
https://github.com/brouwern/compbio4all.git

## Associate working directory with repo via RStudio

Clone into the repo by hand by creating an RStudio Project: New project / New repro / etc

## Set up package with usethis

I then use usethis::create_package() within my project directory to build the basic package infrastructure  This overwrites the initial project (after a handy prompt in the console).

```{r eval = F}
usethis::create_package(path = getwd())
```

### Create vignette infrastructure

I create vignette infrastructure and dummy vignette

```{r, eval = F}
usethis::use_vignette("temp")
```

I then write my vignettes using this template and change the file name.  I use "a)", "b)", "c)" as a prefix on my vignettes so that show up in order when the website is rendered using `packagedown`.


### Copy in package-making scripts

I copy this script and related ones for maintaining the package into the vignettes folder.

### Create "readme" and "news" files



```{r, eval = F}
usethis::use_readme_md()
usethis::use_news_md()
```



### Populating the readme file

The `readme` shows up on the GitHub landing page and so remember to fill out relevant information such as use-cases there.

I add basic information to the readme fill and add sample code for a call to devtools::install_github() to install the package.  The chunk below is an example:

<!-- #### Example materials for readme -->

<!-- The development version of `thispackage` is on GitHub.  If you don't already have it, you will need to install the devtools package -->

<!-- ``` r -->
<!-- install.packages("devtools") -->
<!-- ``` -->

<!-- You can then install the `thispackage` from GitHub with: -->

<!-- ``` r -->
<!-- devtools::install_github("brouwern/compbio4all") -->
<!-- ``` -->



### Create git ignore


Add typical files to gitignore: '.Rproj.user', '.Rhistory', '.Rdata', '.DS_Store'

```{r}
usethis::git_vaccinate()
```


I always have trouble with `gitingore` and haven't taken the time to learn more about it.  I am not sure how to use this properly yet.

```{r, eval = F}
usethis::use_git_ignore(".pdf", directory = ".")
usethis::use_git_ignore(".xls", directory = ".")
usethis::use_git_ignore(".xlsx", directory = ".")
usethis::use_git_ignore(".docx", directory = ".")
```


### Add required packages

These are the packages I typically use.  I should vectorize this so it would look nicer : )

```{r, eval = F}
# data viz
usethis::use_package("ggplot2", "Imports")
usethis::use_package("ggpubr",   "Imports")
#usethis::use_package("scatterplot3d",   "Suggests")
usethis::use_package("GGally",   "Suggests")
usethis::use_package("cowplot",   "Imports")
usethis::use_package("here",   "Imports")

# data cleaning
usethis::use_package("dplyr",   "Imports")
usethis::use_package("tidyr",   "Imports")
usethis::use_package("stringr",   "Imports")

usethis::use_package("ape", "Suggests")
usethis::use_package("phangorn", "Suggests")


usethis::use_package("rentrez", "Imports")
usethis::use_package("seqinr", "Imports")
#usethis::use_package("UniprotR",   "Imports") # not used?
usethis::use_package("Biostrings",   "Imports") # oh boy
usethis::use_package("MASS",   "Imports")


usethis::use_package("pander",   "Imports")
usethis::use_package("bio3d",   "Suggests")

usethis::use_package("survival",   "Suggests")
usethis::use_package("reshape2",   "Imports")

usethis::use_package("rpart",   "Suggests")
# #usethis::use_package("blme",   "Imports")

# docxtractr

usethis::use_package("evd", "Suggests")

usethis::use_package("HistData", "Suggests")
usethis::use_package("flextable", "Suggests")
usethis::use_package("webshot", "Suggests")
usethis::use_package("cluster", "Suggests")


usethis::use_package("iNEXT", "Suggests")
usethis::use_package("gplots", "Suggests")
# devtools::use_package("lawstat", "Imports")
# devtools::use_package("effsize", "Imports")
# devtools::use_package("metafor", "Imports")
# devtools::use_package("Formula", "Imports")
# devtools::use_package("data.tree", "Imports")


```



### Don't save/load user workspace between sessions

```{r, eval = F}
usethis::use_blank_slate()
```


### Use roxygen for documentation

Where would we be without roxygen for writing documentation files?!?

```{r, eval = F}
usethis::use_roxygen_md()
```


### Package-level documents

"Adds a dummy .R file that will prompt roxygen to generate basic package-level documentation. "

```{r, eval = F}
usethis::use_package_doc()
```


### Use pkgdown

For making a handy front end website that displays documentation and vignettes for the package.

```{r, eval = F}
usethis::use_pkgdown()
```



## Set up data

Create folder for external data.

R packages often have a folder called `/inst` which stands for "installed."  This folder is usually for miscellaneous files associated with the package.  This includes external data (`/extdata`) such as `.csv` files.

```{r, eval = F}
usethis::use_directory("/inst")
usethis::use_directory("/inst/extdata")
```



### Look at any data in my extdata file

External data is stored in "/inst/extdata"

```{r, eval = F}
list.files(here::here("/inst/extdata"))
```

#### Raw data

I copy raw data files into the `/inst/extdata`.  

#### Raw data prep

If there are any data processing steps that I don't want to include in the vignettes I put them into the directory `packagename/data-raw`.  This structure for this directory and data prep script can be made using `use_data_raw()`

```{r, eval = F}
usethis::use_data_raw()
```


## License

For good information on licenses see Karl Broman's blog: 
http://kbroman.org/pkg_primer/pages/licenses.html

Plain text versions of licenses can be found at
https://creativecommons.org/2011/04/15/plaintext-versions-of-creative-commons-licenses-and-cc0/

```{r, eval = F}
usethis::use_ccby_license()
```

An error can occur if you have a separate license file.  eg

>File LICENSE is not mentioned in the DESCRIPTION file.

This can be fixed by changing the description from "License: GPL-3" to "License: GPL-3 +  file LICENSE"

http://r-pkgs.had.co.nz/description.html#license

### Spell check

I usually spell check as I go - and I always need it!  You can add spell checking as a unit test using `use_spell_check()`


## Documenting datasets

All data sets must be documented.  

```{r, eval = F}
usethis::use_r(name = "fledgedata")

```

A minimal R dataset .R file looks like this
```{r, eval = F}
#' Full dataset dfor Meillere et al. 2015
"telosful"
```

A standard R dataset .R file looks like this
(https://r-pkgs.org/data.html)

```{r, eval = F}
#' Prices of 50,000 round cut diamonds.
#'
#' A dataset containing the prices and other attributes of almost 54,000
#' diamonds.
#'
#' @format A data frame with 53940 rows and 10 variables:
#' \describe{
#'   \item{price}{price, in US dollars}
#'   \item{carat}{weight of the diamond, in carats}
#'   ...
#' }
#' @source \url{http://www.diamondse.info/}
"diamonds"
```

### To do list

To do once a package is made, the following things needs to be updated

1. Push stuff to git
1. Add info to the `readme`, such as basic use case
1. Add info to license


### Other potentially useful usethis functions

* use_r() Create or edit a .R file
* use_build_ignore() Add files to .Rbuildignore
* use_version() use_dev_version() Increment package version
* edit_r_profile() 
* edit_r_environ() 
* edit_r_makevars() 
* edit_rstudio_snippets() 
* edit_git_config() 
* edit_git_ignore()
