---
title: "Latex equations for biologists from A to Z"
description: |
  From "density dependent growth" to "matrix population model".
author: Nathan L. Brouwer
date: 2022-05-23
output:
  distill::distill_article:
    self_contained: false
draft: true
---





```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

I'm working on a book project called "[Ecology For All!](https://bio.libretexts.org/Courses/Gettysburg_College/01%3A_Ecology_for_All)" and spending a lot of time writing equations in latex.  I write up equations like this maybe once a year and have to review every time what to do, and spend time knocking the code around to get what I want.  

There are many good introductions to Latex and Latex equations out there.  Below, I present just a working list of common equations in biology.

## A very brief primer of Latex equations

There are two ways to present Latex equation: **inline** and **display**.  Inline equations appear in a line of normal text, like this: \(\lambda = \frac{N_{t+1}}{N_t}\).  Display mode equations appear on a separate line:

\[\lambda = \frac{N_{t+1}}{N_t}\]

The only difference between inline and display markdown up is the inline is preceded by `\(` and closed by `\)`, while display mode is preceded by `\[` and closed by `\]`.

In an Rmarkdown document inline equations can therefore be included like this: 

text text text`\(\lambda = \frac{N_{t+1}}{N_t}\)` text text

This will be rendered as

text text text \(\lambda = \frac{N_{t+1}}{N_t}\) text text.

Display equation are written like this:

some text

`\[\lambda = \frac{N_{t+1}}{N_t}\]`

some more text



## Latex equations A to Z

## B

### Beta coefficients for linear regression

Inline: \(y_i = \beta_0 + \beta_1*x_i\)

Inline: `\(y_i = \beta_0 + \\beta_1*x_i\)`

Solo:

\[y_i = \beta_0 + \beta_1*x_i\]

`\[y_i = \beta_0 + \beta_1*x_i\]`

## D

### Density dependent population growth

### Differential equation dN/dT

\[\frac{dN}{dT} = bN - dN = (b - d)N \nonumber\]

\[\frac{dN}{dT} = rN \nonumber\]

## E

### Exponents 

\(N_t=N_0e^{rt}\)

### Exponential growth

\(N_t=N_0e^{rt}\)

## F

### Fractions

Inline basic: \(\frac{x}{y}\)

Inline example: \(\frac{N_{t+1}}{N_t}\)

Solo:

\[\frac{x}{y}\]

\[\frac{N_{t+1}}{N_t}\]

### Formatting

```{r echo = F}
## Formatting
#               #what                 latex           example
exponent  <- c("Exponent",           "`\\(e^x\\)`",         "\\(e^x\\)")
subcript1 <- c("Subscript-simple",   "`\\(N_t\\)`",         "\\(N_t\\)")
subcript2 <- c("Subscript-complex",  "`\\(N_{t+1}\\)`",     "\\(N_{t+1}\\)")
# fraction  <- c("Fraction",           "\frac{x}{y}", "")           # \\(\frac{x}{y}\\)
df <- rbind(exponent , 
            subcript1,
            subcript2
            # fraction
            )
df <- data.frame(df)
rownames(df) <- NULL
names(df) <- c("Format", "Inline Latex", "Example")

pander::pander(df)
```


## G

### Geometric growth

### Greek symbols

Greek symbols are invoked via their name, preceded by a slash.  \(\alpha\) is therefore genereated via ` \(\alpha\) `.  Uppercase, when different than the Enlighs equivalent, are generated by capitalizing the Greek name: \(\Gamma\) is therefore generated via ` \(\Gamma\) `.  When the uppercase Greek and English are the same the markup for normal text within an equation is used: `\\(\\textrm{Z}\\)` yields uppercase Zeta, which is just \(\textrm{Z}\).


```{r, echo = F}
alpha <- c("alpha","\\(\\alpha\\)","\\(\\textrm{A}\\)","`\\(\\alpha\\)`","`\\(\\textrm{A}\\)`")
beta  <- c("beta" ,"\\(\\beta\\)", "\\(\\textrm{B}\\)","`\\(\\beta\\)`", "`\\(\\textrm{B}\\)`")
gamma <- c("gamma","\\(\\gamma\\)","\\(\\Gamma\\)",    "`\\(\\gamma\\)`","`\\(\\Gamma\\)`")
delta <- c("delta","\\(\\delta\\)","\\(\\Delta\\)",    "`\\(\\delta\\)`","`\\(\\Delta\\)`")
epsilon<-c("epsilon","\\(\\epsilon\\)","\\(\\textrm{E}\\)", "`\\(\\epsilon\\)`","`\\(\\textrm{E}\\)`")

zeta <- c("zeta","\\(\\zeta\\)","\\(\\textrm{Z}\\)",    "`\\(\\zeta\\)`","`\\(\\textrm{Z}\\)`")
#eta,theta,iota,kappa,
lambda <-c("lambda","\\(\\lambda\\)","\\(\\textrm{L}\\)", "`\\(\\lambda\\)`","`\\(\\textrm{L}\\)`")
mu <-c("mu","\\(\\mu\\)","\\(\\textrm{M}\\)", "`\\(\\mu\\)`","`\\(\\textrm{M}\\)`")
#nu,xi,omikron,pi, rho,
sigma <- c("sigma","\\(\\sigma\\)","\\(\\Sigma\\)",    "`\\(\\sigma\\)`","`\\(\\Sigma\\)`")
#, tau, upsilon, 
phi  <- c("phi","\\(\\phi\\)","\\(\\Phi\\)",    "`\\(\\phi\\)`","`\\(\\Phi\\)`")
chi <-c("chi","\\(\\chi\\)","\\(\\textrm{X}\\)", "`\\(\\chi\\)`","`\\(\\textrm{X}\\)`")
psi  <- c("psi","\\(\\psi\\)","\\(\\Psi\\)",    "`\\(\\psi\\)`","`\\(\\Psi\\)`")
#, omega

df <- rbind(alpha, beta, gamma,delta,
            epsilon,
            zeta,
            #eta,theta,iota,kappa
            lambda,
            mu,
            #nu,xi,omikron,pi, rho,
sigma, 
#, tau, upsilon, 
phi, 
chi, 
psi#, omega
)
rownames(df) <- NULL
colnames(df) <- c("Name",
                  "lower-case\nsymbol","upper-case\nsymbol",
                  "lower-case\nLatex","upper-case\nLatex")
pander::pander(df, keep.line.breaks = TRUE)


```




## I

### Identity matrix

\[\begin{bmatrix}
1 & 0 \\
0 & 1
\end{bmatrix}\]

\[\boldsymbol{I} = \begin{bmatrix}
1 & 0 \\
0 & 1
\end{bmatrix}\]

Note: the use of `boldsymbol{}` to get the bold I may be outmoded or otherwise problematic.

## L

### Lambda

Inline basic: \(\lambda\)

Inline example: \(\lambda = \frac{N_{t+1}}{N_t}\)

Solo:

\[\lambda\]

\[\lambda = \frac{N_{t+1}}{N_t}\]

### Linear model

### Logistic growth

\[\frac{dN}{dT} = r_{max}\frac{dN}{dT} =r_{max}*N*\frac{K - N}{K} \nonumber\]

## M

### Matrix model

\[\begin{bmatrix}
1 & 2 \\
3 & 4
\end{bmatrix}\]

\[\begin{bmatrix}
1 & 2 & 3\\
4 & 5 & 6\\
7 & 8 & 9
\end{bmatrix}\]


## Subscripts

\(n_x\)

\(N_x\)

\(N_{t_{double}}\)
